

#Read data


To create the big matrix, I pick up the years 2005 and 2008 and create a csv document of all these data.
```{r}
#library(biganalytics)
#X <- read.big.matrix("AirlineData0508.csv", header = TRUE, 
                     #backingfile = "air0508.bin",
                     #descriptorfile = "air0508.desc",
                     #type = "integer", extraCols = "age")
#X05 <- read.big.matrix("2005.csv", header = TRUE, 
                    # backingfile = "air05.bin",
                    # descriptorfile = "air05.desc",
                    # type = "integer", extraCols = "age")
#X08<- read.big.matrix("2008.csv", header = TRUE, 
                    # backingfile = "air08.bin",
                    # descriptorfile = "air08.desc",
                    # type = "integer", extraCols = "age")
```


read the data and the big matrix.
```{r}
library(data.table)
library(biganalytics)
library(foreach)
#x0508<- attach.big.matrix("air0508.desc")
#data08=fread("~/project/2008.csv")
#save(data08, file = 'data08.RDdata')
#data05=fread("~/project/2005.csv")
#save(data05, file = 'data05.RDdata')
load('data05.RDdata')
load('data08.RDdata')
#head(data05)
```

#Check the NA in the data.
Na means we do not have the information about that flight. Since we are  interested in the reason or related variables about the specific type of delay. Even though there are many NAs in the data set of 2008, there still a large number of observations in our data. It is enough to do the further research.
```{r}
#check 2005
na05=sum(is.na(data05[,'WeatherDelay'])|is.na(data05[,'SecurityDelay'])|is.na(data05[,'LateAircraftDelay'])|is.na(data05[,'NASDelay'])|is.na(data05[,'CarrierDelay']))
na05
nrow(data05)-na05


#check 2008
na08=sum(is.na(data08[,'WeatherDelay'])&is.na(data08[,'SecurityDelay'])&is.na(data08[,'LateAircraftDelay'])&is.na(data08[,'NASDelay'])&is.na(data08[,'CarrierDelay']))
na08
nrow(data08)-na08



sum(is.na(data08[,'WeatherDelay']))
sum(is.na(data08[,'SecurityDelay']))
sum(is.na(data08[,'LateAircraftDelay']))
sum(is.na(data08[,'NASDelay']))
sum(is.na(data08[,'CarrierDelay']))
```

#calculate the proportion of the average delay for 2005 data.
```{r}
WeatherDelay05=sum(data05[which(data05[,'WeatherDelay']>0),'WeatherDelay'])/nrow(data05)
SecurityDelay05=sum(data05[which(data05[,'SecurityDelay']>0),'SecurityDelay'])/nrow(data05)
CarrierDelay05=sum(data05[which(data05[,'CarrierDelay']>0),'CarrierDelay'])/nrow(data05)
NASDelay05=sum(data05[which(data05[,'NASDelay']>0),'NASDelay'])/nrow(data05)
LateAircraftDelay05=sum(data05[which(data05[,'LateAircraftDelay']>0),'LateAircraftDelay'])/nrow(data05)
delay05=cbind(LateAircraftDelay05,NASDelay05,CarrierDelay05,WeatherDelay05,SecurityDelay05)
percent05=numeric(5)
for(i in 1:5){
  percent05[i]=delay05[i]/sum(delay05)
}
delayType=c('LateAircraftDelay','NASDelay','CarrierDelay','WeatherDelay','SecurityDelay')
cbind(delayType,percent05)
 delayType <- paste( delayType, round(percent05*100,2),'%')
#colnames(data05)
pie(percent05, labels = delayType, main="Pie Chart of Delay Type for 2005")
```


##calculate the proportion of the average delay for 2008 data.
```{r}
#So I only consider the non-NA part.
#colnames(x08)
#data08new=data08[complete.cases(data08[ , 25:29]),]
data08new<- attach.big.matrix("air08.desc")

WeatherDelay08=sum(data08new[which(data08new[,'WeatherDelay']>0),'WeatherDelay'])/nrow(data08new)
SecurityDelay08=sum(data08new[which(data08new[,'SecurityDelay']>0),'SecurityDelay'])/nrow(data08new)
CarrierDelay08=sum(data08new[which(data08new[,'CarrierDelay']>0),'CarrierDelay'])/nrow(data08new)
NASDelay08=sum(data08new[which(data08new[,'NASDelay']>0),'NASDelay'])/nrow(data08new)
LateAircraftDelay08=sum(data08new[which(data08new[,'LateAircraftDelay']>0),'LateAircraftDelay'])/nrow(data08new)
delay08=cbind(LateAircraftDelay08,NASDelay08,CarrierDelay08,WeatherDelay08,SecurityDelay08)
percent08=numeric(5)
for(i in 1:5){
  percent08[i]=delay08[i]/sum(delay08)
}
delayType=c('LateAircraftDelay','NASDelay','CarrierDelay','WeatherDelay','SecurityDelay')
cbind(delayType,percent08)
 delayType <- paste( delayType, round(percent08*100,2),'%')
pie(percent08, labels = delayType, main="Pie Chart of Delay Type for 2008")
```

#2005
```{r}
arrdelayInd<-which(data05[,"ArrDelay"]>0)
arrdelayed<-data05[arrdelayInd,c('LateAircraftDelay','NASDelay','CarrierDelay','WeatherDelay','SecurityDelay')]
wadelayed<-which(arrdelayed[, "WeatherDelay"]>0)
sadelayed<-which(arrdelayed[, "SecurityDelay"]>0)
nadelayed<-which(arrdelayed[, "NASDelay"]>0)
cadelayed<-which(arrdelayed[, "CarrierDelay"]>0)
ladelayed<-which(arrdelayed[, "LateAircraftDelay"]>0)

# Percentage of late arriving flights that had weather delays
# About 3.6% of late arriving flights had weather delays
length(wadelayed)/nrow(arrdelayed)

# Percentage of late arriving flights that had security delays
# About .21% of late arrivig flights had security delays
length(sadelayed)/nrow(arrdelayed)

# Percentage of late arriving flights that had NASDelay
# About 29% of late arriving flights had NASDelay
length(nadelayed)/nrow(arrdelayed)

# Percentage of late arriving flights that had CarrierDelay
# About 21.5% of late arrivig flights had CarrierDelay
length(cadelayed)/nrow(arrdelayed)

# Percentage of late arriving flights that had LateAircraftDelay
# About 20.5% of late arrivig flights had LateAircraftDelay
length(ladelayed)/nrow(arrdelayed)


# Average weather delay for flights with arrival delays & average weather delay for late arriving flights that had weather delays
# Average weather delay for flights delayed by weather is much higher because most late arriving flights are not delayed by weather.
a=unlist(arrdelayed[,"WeatherDelay"])
c(mean(a,na.rm=TRUE), mean(unlist(arrdelayed[wadelayed,"WeatherDelay"]),na.rm=TRUE))
# Similarly for security delays
c(mean(unlist(arrdelayed[,"SecurityDelay"]),na.rm=TRUE), mean(unlist(arrdelayed[sadelayed,"SecurityDelay"]),na.rm=TRUE))
#LateAircraftDelay
c(mean(unlist(arrdelayed[,"LateAircraftDelay"]),na.rm=TRUE), mean(unlist(arrdelayed[ladelayed,"LateAircraftDelay"]),na.rm=TRUE))
#CarrierDelay
c(mean(unlist(arrdelayed[,"CarrierDelay"]),na.rm=TRUE), mean(unlist(arrdelayed[cadelayed,"CarrierDelay"]),na.rm=TRUE))
c(mean(unlist(arrdelayed[,"NASDelay"]),na.rm=TRUE), mean(unlist(arrdelayed[nadelayed,"NASDelay"]),na.rm=TRUE))


# Average arrival delays for fligths delayed by weather and delayed by security
c(mean(unlist(data05[which(data05[,"WeatherDelay"]>0),"ArrDelay"]),na.rm=TRUE),
  mean(unlist(data05[which(data05[,"SecurityDelay"]>0),"ArrDelay"]),na.rm=TRUE),mean(unlist(data05[which(data05[,"CarrierDelay"]>0),"ArrDelay"]),na.rm=TRUE),mean(unlist(data05[which(data05[,"NASDelay"]>0),"ArrDelay"]),na.rm=TRUE),mean(unlist(data05[which(data05[,"LateAircraftDelay"]>0),"ArrDelay"]),na.rm=TRUE))

# We could note from the examples above that weather delays tend to be longer than security delays on average for flights that 
# arrive late and in general and weather delays happen more often than security delays.
# We can also conclude that, on average, flights delayed by weather tend to have longer arrival delays than those delayed by security issues.

# We could consider flights delayed by both types, but there are very few of these cases.
bothdelays=which(data05[,"SecurityDelay"]>0 & data05[,"WeatherDelay"]>0&data05[,"LateAircraftDelay"]>0&data05[,"NASDelay"]>0&data05[,"SecurityDelay"]>0)
mean(unlist(data05[bothdelays,"ArrDelay"]),na.rm=TRUE)
length(bothdelays)
```

#2008
```{r}
arrdelayInd<-which(data08new[,"ArrDelay"]>0)
arrdelayed<-data08new[arrdelayInd,c('LateAircraftDelay','NASDelay','CarrierDelay','WeatherDelay','SecurityDelay')]
wadelayed<-which(arrdelayed[, "WeatherDelay"]>0)
sadelayed<-which(arrdelayed[, "SecurityDelay"]>0)
nadelayed<-which(arrdelayed[, "NASDelay"]>0)
cadelayed<-which(arrdelayed[, "CarrierDelay"]>0)
ladelayed<-which(arrdelayed[, "LateAircraftDelay"]>0)

# Percentage of late arriving flights that had weather delays
# About 3.6% of late arriving flights had weather delays
length(wadelayed)/nrow(arrdelayed)

# Percentage of late arriving flights that had security delays
# About .21% of late arrivig flights had security delays
length(sadelayed)/nrow(arrdelayed)

# Percentage of late arriving flights that had NASDelay
# About 29% of late arriving flights had NASDelay
length(nadelayed)/nrow(arrdelayed)

# Percentage of late arriving flights that had CarrierDelay
# About 21.5% of late arrivig flights had CarrierDelay
length(cadelayed)/nrow(arrdelayed)

# Percentage of late arriving flights that had LateAircraftDelay
# About 20.5% of late arrivig flights had LateAircraftDelay
length(ladelayed)/nrow(arrdelayed)


# Average weather delay for flights with arrival delays & average weather delay for late arriving flights that had weather delays
# Average weather delay for flights delayed by weather is much higher because most late arriving flights are not delayed by weather.
a=unlist(arrdelayed[,"WeatherDelay"])
c(mean(a,na.rm=TRUE), mean(unlist(arrdelayed[wadelayed,"WeatherDelay"]),na.rm=TRUE))
# Similarly for security delays
c(mean(unlist(arrdelayed[,"SecurityDelay"]),na.rm=TRUE), mean(unlist(arrdelayed[sadelayed,"SecurityDelay"]),na.rm=TRUE))
#LateAircraftDelay
c(mean(unlist(arrdelayed[,"LateAircraftDelay"]),na.rm=TRUE), mean(unlist(arrdelayed[ladelayed,"LateAircraftDelay"]),na.rm=TRUE))
#CarrierDelay
c(mean(unlist(arrdelayed[,"CarrierDelay"]),na.rm=TRUE), mean(unlist(arrdelayed[cadelayed,"CarrierDelay"]),na.rm=TRUE))
c(mean(unlist(arrdelayed[,"NASDelay"]),na.rm=TRUE), mean(unlist(arrdelayed[nadelayed,"NASDelay"]),na.rm=TRUE))


# Average arrival delays for fligths delayed by weather and delayed by security
c(mean(unlist(data08new[which(data08new[,"WeatherDelay"]>0),"ArrDelay"]),na.rm=TRUE),
  mean(unlist(data08new[which(data08new[,"SecurityDelay"]>0),"ArrDelay"]),na.rm=TRUE),mean(unlist(data08new[which(data08new[,"CarrierDelay"]>0),"ArrDelay"]),na.rm=TRUE),mean(unlist(data08new[which(data08new[,"NASDelay"]>0),"ArrDelay"]),na.rm=TRUE),mean(unlist(data08new[which(data08new[,"LateAircraftDelay"]>0),"ArrDelay"]),na.rm=TRUE))

# We could note from the examples above that weather delays tend to be longer than security delays on average for flights that 
# arrive late and in general and weather delays happen more often than security delays.
# We can also conclude that, on average, flights delayed by weather tend to have longer arrival delays than those delayed by security issues.

# We could consider flights delayed by both types, but there are very few of these cases.
bothdelays=which(data08new[,"SecurityDelay"]>0 & data08new[,"WeatherDelay"]>0&data08new[,"LateAircraftDelay"]>0&data08new[,"NASDelay"]>0&data08new[,"SecurityDelay"]>0)
mean(unlist(data08new[bothdelays,"ArrDelay"]),na.rm=TRUE)
length(bothdelays)
```
```{r}
library(biganalytics)
x05<- attach.big.matrix("air05.desc")
x08<- attach.big.matrix("air08.desc")
blm3 <- biglm.big.matrix( ArrDelay ~ WeatherDelay+SecurityDelay+LateAircraftDelay+NASDelay+SecurityDelay, data = x05)
summary(blm3)$rsq
blm3 <- biglm.big.matrix( ArrDelay ~LateAircraftDelay+NASDelay+SecurityDelay, data = x05)
summary(blm3)$rsq


blm3 <- biglm.big.matrix( ArrDelay ~ WeatherDelay+SecurityDelay+LateAircraftDelay+NASDelay+SecurityDelay, data = x08)
summary(blm3)$rsq

blm3 <- biglm.big.matrix( ArrDelay ~LateAircraftDelay+NASDelay+SecurityDelay, data = x08)
summary(blm3)$rsq
```
#So we focus on LateAircraftDelay.
To explore the related variable behind the specific type of delay, we could ingore the flight with value 0 in that type of delay.
```{r}
library(biganalytics)
x05<- attach.big.matrix("air05.desc")
x08<- attach.big.matrix("air08.desc")
blm3 <- biglm.big.matrix( ArrDelay ~ WeatherDelay+SecurityDelay+LateAircraftDelay+NASDelay+SecurityDelay, data = x05)
summary(blm3)$rsq
blm3 <- biglm.big.matrix( ArrDelay ~LateAircraftDelay+NASDelay+SecurityDelay, data = x05)
summary(blm3)$rsq


blm3 <- biglm.big.matrix( ArrDelay ~ WeatherDelay+SecurityDelay+LateAircraftDelay+NASDelay+SecurityDelay, data = x08)
summary(blm3)$rsq

blm3 <- biglm.big.matrix( ArrDelay ~LateAircraftDelay+NASDelay+SecurityDelay, data = x08)
summary(blm3)$rsq
```
#So we focus on LateAircraftDelay.
#rate
```{r}
library(biganalytics)
library(foreach)
x05<- attach.big.matrix("air05.desc")
x05[which(x05[,'LateAircraftDelay']>0),'LateAircraftDelay']=1
x05[which((x05[,'LateAircraftDelay']<=0)),'LateAircraftDelay']=0
# Divide CRSDepTime by 100 and take the floor to 
# get the departure hour.
depHours <- floor(x05[,"CRSDepTime"]/100)
# Set the departure hours listed as 24 to 0.
depHours[depHours==24] <- 0
# Split on the hours.
hourInds <- split(1:length(depHours), depHours)
# Create a variable to hold the quantile probabilities.
#myProbs <- c(0.55)
# Use foreach to find the quantiles for each hour.
LateAircraftDelayQuantiles05 <- foreach( hour = hourInds, .combine=cbind) %do% {
  require(bigmemory)
  mean(x05[hour, "LateAircraftDelay"], na.rm = TRUE)
}
# Clean up the column names.
colnames(LateAircraftDelayQuantiles05) <- names(hourInds)
```

```{r}
x08<- attach.big.matrix("air08.desc")
x08[which((x08[,'LateAircraftDelay']>0)),'LateAircraftDelay']=1
x08[which((x08[,'LateAircraftDelay']<=0)),'LateAircraftDelay']=0
nrow(x08)
head(x08)
#is.na(x08[1,'CancellationCode'])
#x08=x08[which(is.na(x08[,'LateAircraftDelay'])),]
#sum(is.na(x08[,'LateAircraftDelay']))

# Divide CRSDepTime by 100 and take the floor to 
# get the departure hour.
depHours <- floor(x08[,"CRSDepTime"]/100)
# Set the departure hours listed as 24 to 0.
depHours[depHours==24] <- 0
# Split on the hours.
hourInds <- split(1:length(depHours), depHours)
# Create a variable to hold the quantile probabilities.
myProbs <- c(0.55)
# Use foreach to find the quantiles for each hour.
LateAircraftDelayQuantiles08 <- foreach( hour = hourInds, .combine=cbind) %do% {
  require(bigmemory)
 mean(x08[hour, "LateAircraftDelay"], na.rm = TRUE)
}

# Clean up the column names.
colnames(LateAircraftDelayQuantiles08) <- names(hourInds)
# See delay quantiles for the data.
library(ggplot2)
library(reshape2) # load library for melt; this is not in the text code
dq08 <- melt(LateAircraftDelayQuantiles08)
dq05=melt(LateAircraftDelayQuantiles05)

names(dq05) <- c("percentile", "hour", "delay05")
names(dq08) <- c("percentile", "hour", "delay08")

# Line plot with multiple groups
ggplot() +
  geom_line(data=dq05, aes(x=hour, y=delay05),color="red",size=1.2)+geom_line(data=dq08, aes(x=hour, y=delay08), color="blue", size=1.2)+labs(title = " Late Aircraft Delay rate vs  Departure Hour")+ ylab("Late Aircraft Delay rate")+xlab('Hour')+
  scale_color_discrete(name ="Year", labels = c("2005", "2008")) #??????
```
```{r}
library(biganalytics)
library(foreach)
x05<- attach.big.matrix("air05.desc")
x05=x05[which(x05[,'LateAircraftDelay']>0),]
# Divide CRSDepTime by 100 and take the floor to 
# get the departure hour.
depHours <- floor(x05[,"CRSDepTime"]/100)
# Set the departure hours listed as 24 to 0.
depHours[depHours==24] <- 0
# Split on the hours.
hourInds <- split(1:length(depHours), depHours)
# Create a variable to hold the quantile probabilities.
myProbs <- c(0.55)
# Use foreach to find the quantiles for each hour.
LateAircraftDelayQuantiles05 <- foreach( hour = hourInds, .combine=cbind) %do% {
  require(bigmemory)
  quantile(x05[hour, "LateAircraftDelay"], myProbs, na.rm = TRUE)
}
# Clean up the column names.
colnames(LateAircraftDelayQuantiles05) <- names(hourInds)
```

```{r}
x08<- attach.big.matrix("air08.desc")
x08=x08[which((x08[,'LateAircraftDelay']>0)),]
nrow(x08)
head(x08)
#is.na(x08[1,'CancellationCode'])
#x08=x08[which(is.na(x08[,'LateAircraftDelay'])),]
#sum(is.na(x08[,'LateAircraftDelay']))

# Divide CRSDepTime by 100 and take the floor to 
# get the departure hour.
depHours <- floor(x08[,"CRSDepTime"]/100)
# Set the departure hours listed as 24 to 0.
depHours[depHours==24] <- 0
# Split on the hours.
hourInds <- split(1:length(depHours), depHours)
# Create a variable to hold the quantile probabilities.
myProbs <- c(0.55)
# Use foreach to find the quantiles for each hour.
LateAircraftDelayQuantiles08 <- foreach( hour = hourInds, .combine=cbind) %do% {
  require(bigmemory)
  quantile(x08[hour, "DepDelay"], myProbs, na.rm = TRUE)
}

# Clean up the column names.
colnames(LateAircraftDelayQuantiles08) <- names(hourInds)
# See delay quantiles for the data.
library(ggplot2)
library(reshape2) # load library for melt; this is not in the text code
dq08 <- melt(LateAircraftDelayQuantiles08)
dq05=melt(LateAircraftDelayQuantiles05)

names(dq05) <- c("percentile", "hour", "delay05")
names(dq08) <- c("percentile", "hour", "delay08")

# Line plot with multiple groups
ggplot() +
  geom_line(data=dq05, aes(x=hour, y=delay05),color="red",size=1.2)+geom_line(data=dq08, aes(x=hour, y=delay08), color="blue", size=1.2)+labs(title = " Late Aircraft Delay vs  Departure Hour")+ ylab("Late Aircraft Delay")+xlab('Hour')+
  scale_color_discrete(name ="Year", labels = c("2005", "2008")) #??????
```



#Merge data.

```{r}
#for 2005
airport <- read.csv("airports.csv")
destination_airport <- read.csv("airports.csv")
names(destination_airport) <- c("destination_iata","destination_airport", "destination_city","destination_state", "destination_country", "destination_lat", "destination_long")

#merge origin airports
data_all05<- merge(data05, airport, by.x = 'Origin', by.y = 'iata')
#merge destination airports
data_all05 <- merge (data_all05, destination_airport,  by.x = 'Dest', by.y = 'destination_iata')
#merge plane
plane <- read.csv("plane-data.csv")
colnames(plane[c("year")]) <- "plane_year"
data_all05 <- merge (data_all05, plane ,by.x = 'TailNum', by.y = 'tailnum', all.x = T)
#merge carrier
carrier=read.csv("carriers.csv")
data_all05 <- merge (data_all05, carrier,by.x = 'UniqueCarrier', by.y = 'Code', all.x = T)
head(data_all05)
nrow(data_all05)
ncol(data_all05)

#2008
#merge origin airports
data_all08<- merge(data08, airport, by.x = 'Origin', by.y = 'iata')
#merge destination airports
data_all08 <- merge (data_all08, destination_airport,  by.x = 'Dest', by.y = 'destination_iata')
#merge plane
data_all08 <- merge (data_all08, plane ,by.x = 'TailNum', by.y = 'tailnum', all.x = T)
#merge carrier
data_all08 <- merge (data_all08, carrier,by.x = 'UniqueCarrier', by.y = 'Code', all.x = T)
head(data_all08)
nrow(data_all08)
ncol(data_all08)
```

```{r}
#save(data_all08, file = 'data_all08.RDdata')
load('data_all08.RDdata')
#save(data_all05, file = 'data_all05.RDdata')
load('data_all05.RDdata')
```


#explore the related variables of NASDelay.

```{r,echo=FALSE,message=FALSE, warning=FALSE, results='hide'}
installIfNeeded = function(cliblist){
  libsNeeded = cliblist
  libsNeeded = libsNeeded[!(libsNeeded %in% installed.packages()[,"Package"])]
  if(length(libsNeeded)>0) install.packages(libsNeeded)
}

installIfNeeded("treemap")

# load the treemap package
library('treemap')

# Any package that is required by the script below is given here:
inst_pkgs = load_pkgs =  c("ggplot2","ggplot2movies", "dplyr","babynames","data.table","Rcpp")
inst_pkgs = inst_pkgs[!(inst_pkgs %in% installed.packages()[,"Package"])]
if(length(inst_pkgs)) install.packages(inst_pkgs)

git_pkgs = git_pkgs_load = c("streamgraph","DT")

git_pkgs = git_pkgs[!(git_pkgs %in% installed.packages()[,"Package"])]

if(length(git_pkgs)){
  library(devtools)
  install_github('rstudio/DT')
  install_github('hrbrmstr/streamgraph')
}

load_pkgs = c(load_pkgs, git_pkgs_load)

# Dynamically load packages
pkgs_loaded = lapply(load_pkgs, require, character.only=T)

```



```{r}
#2005

data_NA05=data_all05[which(data_all05[,'NASDelay']>0),]
#which(data_all05[,'NASDelay']<0)????
nrow(data_NA05)
ncol(data_NA05)
```


```{r}
#average delay time for each origin airport(departure)
airport_NA05_mean <- aggregate(NASDelay ~Origin,data_NA05, mean)
airport_NA05_mean<- airport_NA05_mean[order(airport_NA05_mean$NASDelay),]
airport_NA05_mean
summary(airport_NA05_mean)
treemap(airport_NA05_mean,
        index=c("Origin"),
        vSize="NASDelay",
        vColor="NASDelay",
        type="value",title = 'Average NASDelay for Each Origin Airport in 2005')
```

```{r}
#2008
data_NA08=data_all08[which(data_all08[,'NASDelay']>0),]
data_NA08=data_NA08[which(!is.na(data_NA08[,'NASDelay'])),]
nrow(data_NA08)
ncol(data_NA08)
```

```{r}
#average delay time for each origin airport(departure)
Airport_NA08_mean <- aggregate(NASDelay ~Origin,data_NA08, mean)
Airport_NA08_mean<- Airport_NA08_mean[order(Airport_NA08_mean$NASDelay),]
Airport_NA08_mean
summary(Airport_NA08_mean)

treemap(Airport_NA08_mean,
        index=c("Origin"),
        vSize="NASDelay",
        vColor="NASDelay",
        type="value",title = 'Average NASDelay for Each Origin Airport in 2008')
```
#Explore the NA data
```{r}
#2008
data_NA08NEW=data_all08[which(is.na(data_all08[,'NASDelay'])),]
head(data_NA08NEW)
data_NA08NEW[,'NASDelay']=1
```

```{r}
Airport_NA08_sum<- aggregate(NASDelay ~Origin,data_NA08NEW, sum)
Airport_NA08_sum<- Airport_NA08_sum[order(Airport_NA08_sum$NASDelay),]
Airport_NA08_sum
summary(Airport_NA08_sum)

treemap(Airport_NA08_sum,
        index=c("Origin"),
        vSize="NASDelay",
        vColor="NASDelay",
        type="value",title = 'Count of NA in NASDelay for Each Origin Airport in 2008')
```

#Explore the related variable of CarrierDelay
```{r}
#2005
data_Carrier05=data_all05[which(data_all05[,'CarrierDelay']>0),]
#which(data_all05[,'NASDelay']<0)????
nrow(data_Carrier05)
ncol(data_Carrier05)


#average delay time for each Carrier 
Carrier05_mean <- aggregate(CarrierDelay ~Description,data_Carrier05, mean)
Carrier05_mean<- Carrier05_mean[order(Carrier05_mean$CarrierDelay),]
Carrier05_mean
summary(Carrier05_mean)
treemap(Carrier05_mean,
        index=c("Description"),
        vSize="CarrierDelay",
        vColor="CarrierDelay",
        type="value",title = 'Average CarrierDelay for Each Carrier in 2005')
```

```{r}
#2008
data_Carrier08=data_all08[which(!is.na(data_all08[,'CarrierDelay'])),]
data_Carrier08=data_all08[which(data_Carrier08[,'CarrierDelay']>0),]
#which(data_all05[,'NASDelay']<0)????
nrow(data_Carrier08)
ncol(data_Carrier08)
#names(data_Carrier08)
#head(data_Carrier08)
#average delay time for each Carrier 
Carrier08_mean <- aggregate(CarrierDelay ~Description,data_Carrier08, mean)
Carrier08_mean<- Carrier08_mean[order(Carrier08_mean$CarrierDelay),]
Carrier08_mean
summary(Carrier08_mean)
treemap(Carrier08_mean,
        index=c("Description"),
        vSize="CarrierDelay",
        vColor="CarrierDelay",
        type="value",title = 'Average CarrierDelay for Each Carrier in 2008')
```
#Deal with NA
```{r}
#2008
data_Carrier08NEW=data_all08[which(is.na(data_all08[,'CarrierDelay'])),]
data_Carrier08NEW[,'CarrierDelay']=1

Carrier08_sum<- aggregate(CarrierDelay ~Description,data_Carrier08NEW, sum)
treemap(Carrier08_sum,
 index=c("Description"),
        vSize="CarrierDelay",
        vColor="CarrierDelay",
        type="value",title = 'Count of NA in CarrierDelay for Each Carrier in 2008')

names(data08)
sum(is.na(data08[,'ArrDelay']))
sum(is.na(data08[,'DepDelay']))
```



#find the rountes has the most delays

```{r}
inst_pkgs = load_pkgs =  c("streamR","ROAuth","maps","map_data",
                           "ggplot2", "grid","scales")
inst_pkgs = inst_pkgs[!(inst_pkgs %in% installed.packages()[,"Package"])]
if(length(inst_pkgs)) install.packages(inst_pkgs)

# Dynamically load packages
pkgs_loaded = lapply(load_pkgs, require, character.only=T)
# Dynamically load packages
pkgs_loaded = lapply(load_pkgs, require, character.only=T)
```

```{r}
library( dplyr)
plot.tweets = function(map, pointsa, pointsd){
  x=rbind(pointsa[1,1],pointsd[1,1])
y=rbind(pointsa[1,2],pointsd[1,2])
s=cbind(x,y)
  # Initialize ggplot object with map data
  mapd=ggplot(map) +
    # Add the map data
    geom_map(aes(map_id = region), map = map, fill = "white", color = "grey20", size = 0.25) + 
    # Set map limits
    expand_limits(x = map$long, y = map$lat) + 
    # Add a theme  
    theme(axis.line = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), 
          axis.title = element_blank(), panel.background = element_blank(), panel.border = element_blank(), 
          panel.grid.major = element_blank(), plot.background = element_blank(), 
          plot.margin = unit(0 * c(-1.5, -1.5, -1.5, -1.5), "lines")) + 
    
    # Add points
    geom_point(data = pointsa, aes(x = x, y = y), size = 3, alpha = 1/5, color ='blue')+ 
    geom_point(data = pointsd, aes(x = x, y = y), size = 3, alpha = 1/5, color ='red')+
    geom_line(data=s,aes(x=x,y=y),color='red')
  
  
  # Graph ggplot2 object
  print(mapd)
}

```



```{r}
data05case=data_all05[which(data_all05[,'ArrDelay']>0),]
data08case=data_all08[which(data_all08[,'ArrDelay']>0&!is.na(data_all08[,'ArrDelay'])),]
names(data05case)
```

```{r}
#2005
mean05 <- aggregate(ArrDelay~Origin+long+lat+Dest+destination_airport+destination_long+destination_lat+destination_state+destination_state+destination_country,data05case, mean)
mean05_sort <- mean05[order(-mean05$ArrDelay),]
mean05_sort_10=mean05_sort[1:10,]
#2008
mean08 <- aggregate(ArrDelay~Origin+long++lat+Dest+destination_airport+destination_long+destination_lat+destination_state+destination_state+destination_country,data08case, mean)
mean08_sort <- mean08[order(-mean08$ArrDelay),]
mean08_sort_10=mean08_sort[1:10,]
```

```{r}
#2005
apoints = data.frame(x = as.numeric(mean05_sort_10$destination_long), y = as.numeric(mean05_sort_10$destination_lat))
dpoints = data.frame(x = as.numeric(mean05_sort_10$long), y = as.numeric(mean05_sort_10$lat))
#write.csv(apoints,'a05.csv')
#write.csv(dpoints,'d05.csv')
```


```{r}
usa.map.data = map_data("usa")
# Calling the function to plot the data
plot.tweets(usa.map.data, apoints,dpoints)
```

```{r}
#2008
apoints = data.frame(x = as.numeric(mean08_sort_10$destination_long), y = as.numeric(mean08_sort_10$destination_lat))
dpoints = data.frame(x = as.numeric(mean08_sort_10$long), y = as.numeric(mean08_sort_10$lat))
#write.csv(apoints,'a08.csv')
#write.csv(dpoints,'d08.csv')
```


```{r}
usa.map.data = map_data("usa")
# Calling the function to plot the data
plot.tweets(usa.map.data, apoints,dpoints)
```

